Case-based Arabic Morphological Analysis

### 2010-09-07 tex/sarf.tex
added hadith example figure

%
% File naaclhlt2010.tex
%
% Contact: nasmith@cs.cmu.edu

@@

\usepackage{utf8}
\setarab
\fullvocalize
%\transtrue
\arabtrue

\newcommand{\CharCodeIn}[1]{`\CodeIn{#1}'}
\newcommand{\CodeIn}[1]{{\small\texttt{#1}}}

@@

%\title{Instructions for NAACL HLT 2010 Proceedings\Thanks{This...}}
\title{Case based Arabic Morphological Analysis with 
    Compositional Non-deterministic Automata }

@@

Automated analysis of Arabic data sets, including texts, 
publications, records and digital media is essential
with the huge digital Arabic content available nowadays. 
Morphological analysis is key in current automated 
analysis techniques for Arabic text. 
Current morphological analyzers~\cite{Sughaiyer:04}
use concatenative analysis
to analyze an Arabic word and
consider its internal structure composed of several
{\em morphemes}. A morpheme can be a {\em stem}, or an {\em affix}.
An affix can be a {\em prefix, suffix, } or an {\em infix}.
The analysis of one word may lead to several possible
morphological solutions.
\vocalize
For instance, the word \RL{'a.hmadH} 
may have two valid morphological analyses. 
The letter \RL{'a} may be prefix and the word means ``I praise him''.
The letter \RL{'a} may also be part of the stem 
\RL{'a.hmad} (a proper noun)
and the word means ``his Ahmad''.

\novocalize
The accuracy of the solutions suffer due to inherent difficulties
of morphological analysis of the Arabic language. 
For example, it is common practice to write Arabic text
without short vowels. 
This greatly increases the ambiguity of Arabic text. 
Another difficulty is that  Arabic letters can have up to 
four different forms
corresponding to their position in a word, i.e, beginning,
middle, end of word forms and separate forms. 
This allows the phrase 
\RL{almdrsT}\RL{al_A}
to be visually recognizable
as two separate words \RL{al_A} and \RL{almdrsT} without 
the need of a delimiter such as a space in between. 
This happens because the first word \RL{almdrsT} ends with
\RL{T} a non-connecting letter. 
These words occur often in text and greatly increase the
difficulty of tokenization and are referred to as 
``run-on'' words~\cite{Buckwalter:04}.

@@

We hypothesize that many NLP case studies need the 
morphological analyzer to answer simple queries that do not need 
high complexity and accuracy at the low morphological analysis level 
and that can often compensate at higher level for tolerable inaccuracy at low levels. 
For example, if a query concerns proper names and the 
prefix in question in the analysis connects only to a verb, 
we do not need to further analyze the rest of the word 
to provide an answer.
We find evidence to our hypothesis in~\cite{Maamouri:10} where the 
addition of a new corpus to the Arabic Tree Bank~\cite{Maamouri:04}
with features demanding resolution at an abstraction level
higher than the text itself
led to a refined analyzer~\cite{Kulick:10}.  
We also find evidence in~\cite{Habash:06} that different types of 
morphological analyses behave differently on the same case study. 
We strongly believe and we find evidence in our results that a 
case-based morphological analyzer that allows a case-based
controller to intervene at every decision to 
guide, use, prune, and refine the morphological analysis
is of high utility.

@@

@@

\begin{figure*}[tb]
\center{
\resizebox{1.7\columnwidth}{!}
{ \input{psfigs/FSM.pstex_t}}
\caption{Example composition of the prefix, stem and suffix FSAs.}
\label{f:example}
}
\end{figure*}
% the FSMs for the three words

>>>>

% File naaclhlt2010.tex
%
% Contact: nasmith@cs.cmu.edu

@@

\usepackage{utf8}
\setarab
\fullvocalize
\transtrue
\arabtrue

\newcommand{\CharCodeIn}[1]{`\CodeIn{#1}'}
\newcommand{\CodeIn}[1]{{\small\texttt{#1}}}
\newcommand{\frl}[1]{\fbox{\RL{#1}}}

@@

%\title{Instructions for NAACL HLT 2010 Proceedings\Thanks{This...}}
\title{Case-based Arabic Morphological Analysis }

@@

Automated analysis of Arabic data sets, including texts, 
publications, records and digital media is essential
with the huge digital Arabic content available nowadays. 
Morphological analysis is key in current automated 
analysis techniques for Arabic text. 
Current morphological analyzers~\cite{Sughaiyer:04}
use concatenative analysis
to analyze an Arabic word and
consider its internal structure composed of several
{\em morphemes}. A morpheme can be a {\em stem}, or an {\em affix}.
An affix can be a {\em prefix, suffix, } or an {\em infix}.
The analysis of one word may lead to several possible
morphological solutions.
\vocalize
For instance, the word \RL{'a.hmadH}~\footnote{In this document, we use the default ArabTeX transliteration style ZDMG.}
may have two valid morphological analyses. 
The letter \RL{'a} may be prefix and the word means ``I praise him''.
The letter \RL{'a} may also be part of the stem 
\RL{'a.hmad} (a proper noun)
and the word means ``his Ahmad''.

\novocalize
The accuracy of the solutions suffer due to inherent difficulties
of morphological analysis of the Arabic language. 
For example, it is common practice to write Arabic text
without short vowels. 
This greatly increases the ambiguity of Arabic text. 
Another difficulty is that  Arabic letters can have up to 
four different forms
corresponding to their position in a word, i.e, beginning,
middle, end of word forms and separate forms. 
This allows the phrase \transfalse
\RL{il_A\nospace almdrsT} \transtrue
to be visually recognizable
as two separate words \RL{il_A} (to) and \RL{almdrsT} (the school) 
without the need of a delimiter such as a space in between. 
This happens because the first word \RL{almdrsT} ends with
\RL{T} a non-connecting letter. 
These words occur often in text and greatly increase the
difficulty of tokenization and are referred to as 
``run-on'' words~\cite{Buckwalter:04}.

@@

We hypothesize that many NLP case studies need the 
morphological analyzer to answer queries that do not need 
high accuracy at the low morphological analysis level 
and that can often compensate at higher levels for tolerable inaccuracy 
at low levels. 
For example, if a query concerns proper names and the 
prefix in question in the analysis connects only to a verb, 
we do not need to further analyze the rest of the word 
to provide an answer.
We find evidence to our hypothesis in~\cite{Maamouri:10} where the 
addition of a new corpus to the Arabic Tree Bank~\cite{Maamouri:04}
with features demanding resolution at an abstraction level
higher than the text itself
led to a refined analyzer~\cite{Kulick:10}.  
We also find evidence in~\cite{Habash:06} that different types of 
morphological analyses behave differently on the same case study. 
We strongly believe and we find evidence in our results that a 
case-based morphological analyzer that allows a case-based
controller to intervene at every decision to 
guide, use, prune, and refine the morphological analysis
is of high utility.

@@

\transfalse
\begin{figure}[tb]
\center{
\resizebox{.9\columnwidth}{!}
{ \input{psfigs/exhadith.pstex_t}}
\caption{Hadith abstraction example.}
\label{f:exhadith}
}
\end{figure}
% the FSMs for the three words
\transtrue

@@

\transfalse
\begin{figure*}[tb]
\center{
\resizebox{1.7\columnwidth}{!}
{ \input{psfigs/FSM.pstex_t}}
\caption{Example composition of the prefix, stem and suffix FSAs.}
\label{f:example}
}
\end{figure*}
% the FSMs for the three words
\transtrue
